Watch Out for the Bully! Study of Interference Between Jobs Running Concurrently on Dragonfly Network

### 2016-03-14 draft.tex
update figs in section 5, add conclusion section

\subsection{Simulation Tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability\cite{codes}.  CODES support dragonfly network with high fidelity flit-level simulation\cite{codes-dragonfly}. It has this network workload component that is capable of taking real MPI application traces generated by SST DUMPI\cite{sst} for trace-driven simulations.

@@

In our experiments, the workload running on the dragonfly network under 6 different job placement and routine scheme configurations as listed in Table \ref{tab: placement routing config}. We focus on the analysis of network status while the workload is running, as well as the performance of each individual application. We collected the following result data from the experiments:
\begin{itemize}
    \item router terminal link traffic and busy time.
    \item router local channel traffic and busy time.
    \item router global channel traffic and busy time.
    \item The time application spend on communication. 
\end{itemize}

\TODO{need a figure show workload communication time!!!}

@@

Random placement can randomly allocate the MPI ranks to the physical nodes all over the network, thus prevent the traffic from congesting certain part of the network. Shown as the solid lines in Figure \ref{fig:global-channel-traffic}, the traffic go through global channels are quite uniformly distributed across all the routers in the network, thus no particular global channel will suffer high traffic burden. When the random placement coupled with minimal routing, the packets can avoid unnecessary intermediate forwarding as introduced by adaptive and progressive routing, which results in generating less traffic. That's why the blue solid line is lower than the red and green solid line, shown in figure\ref{fig:global-channel-traffic}. Since the random placement can uniformly distribute the workload traffic, there are no sudden surge of busy time of global channel over the network, shown as solid lines in Fgiure \ref{fig:global-channel-stime}.

@@

However, AMG's performance is quite an exception. As shown in Figure \ref{fig:amg-commtime}, the communication time of AMG is skyrocketing when it running under random placement coupled with (progressive) adaptive routing. The contiguous placement coupled with (progressive) adaptive routing can reach the best performance for AMG. The performance of AMG running under contiguous placement coupled with minimal routing is no better than random coupled with minimal routing, but still much better than random with (progressive) adaptive routing.

% Show the evidence from network level, explain why AMG is different. 
We look into network level to identify the major culprit behind AMG's abnormal behavior with random placement and adaptive routing. By identifying the terminal nodes and routers that each application's traffic go through, we can see that the traffic amount go through the routers that belongs to AMG is skyrocketing when it running with random placement and adaptive routing, shown in Figure \ref{fig:amg-lc-traffic} and Figure \ref{fig:amg-gc-traffic}. Adaptive routing will redirect the traffic of other applications to go through the routers belongs to AMG, and increase the burden of those routers, thus slow down the communication of AMG. \textbf{More explanation about contiguous placement and minimal routing}

CrystalRouter and MultiGrid can get obvious benefit from random placement and adaptive routing, as we can see from Figure \ref{fig:cr-gc-traffic} \ref{fig:cr-lc-traffic} and Figure \ref{fig:mg-gc-traffic} \ref{fig:mg-lc-traffic}, random placement and adaptive routing can redirect their traffic to other region of network, to avoid local congestion as appeared with contiguous and minimal routing. \textbf{More explanation about contiguous placement and minimal routing}

@@

Why AMG is the only victim in the workload? The reason for CrystalRouter and MultiGrid get improvement in the sacrifice of AMG's performance is that AMG is the lest communication-intensive application in the workload. As shown in Figure \ref{fig:3app-data-amount}, AMG has the least amount of data transfer between its ranks.

\begin{figure*}[t!]
  \centering
  \includegraphics[height=2.2in]{data_amount}
  \caption{Applications Data Transfer Amount. AMG has the least amount of data transfer, compared with CR and MG.}
  \label{fig:3app-data-amount}
\end{figure*}

Apparently, random placement and adaptive routing is quite efficient for orchestrate the traffic load in the network. Random placement and adaptive routing can uniformly distribute the workload traffic all over the network, make the load-balanced and avoid hot-spots. However, application like AMG will be impaired in this orchestration. The routers through which the less communication-intensive application traffic used to route, now have more traffic go through them. These extra traffic are from other communication-intensive applications. Beneath the facade of harmony, less communication-intensive application are ``bullied" by their communication-intensive peers in the workload.

We have tried three different congestion sensing schemes that used in adaptive routing\cite{dally-dragonfly}, although there are some variations of the experiments, none of the congestion sensing scheme can prevent the ``bully" from happening.

@@

To validate our assumption, we generate synthetic application, sAMG, which are same the with AMG's original traces in every aspect expect that the data transfer amount 100x greater than AMG. We run the new workload that consists of sAMG, MG and CR on the same dragonfly network with the same placement and routing configurations. As shown in Figure \ref{fig:syn-3app-data-amount}, sAMG become the most communication-intensive application in the new workload.

\begin{figure*}[t!]
  \centering
  \includegraphics[height=2.2in]{syn-wkld/data_amount}
  \caption{Applications Data Transfer Amount. sAMG has the most amount of data transfer, compared with CR and MG.}
  \label{fig:syn-3app-data-amount}
\end{figure*}

The communication time of each application shown in Figure \ref{fig:syn-apps-commtime}. The ``bully" become the ``bullyee". The less communication-intensive application, CrystalRouter and MultiGrid, suffer prolonged communication time when running concurrently with sAMG with random placement and adaptive routing configuration. The routers that belongs to MultiGrid  and now CrystalRouter has extra traffic go through them, shown in Figure  \ref{fig:syn-mg-lc-traffic} \ref{fig:syn-mg-gc-traffic} and Figure \ref{fig:syn-cr-lc-traffic} \ref{fig:syn-cr-gc-traffic}. The extra traffic come from sAMG, that been redirected by adaptive routing. \textbf{More explanation about contiguous placement and minimal routing}.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/commtime}
        \caption{sAMG Communication Time}
        \label{fig:samg-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/commtime}
        \caption{MG Communication Time}
        \label{fig:syn-mg-commtime}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/commtime}
        \caption{CR Communication Time}
        \label{fig:syn-cr-commtime}
    \end{subfigure}%
   \caption{Application communication time. Three applications running concurrently on dragonfly network with different placement and routing configurations. Random placement and adaptive routing can improves AMG's communication time. MG and CR communication time are prolonged with random placement and adaptive routing.}
   \label{fig:syn-apps-commtime}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:samg-lc-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:syn-mg-lc-traffic}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:syn-cr-lc-traffic}
    \end{subfigure}%
   \caption{Each Application Router Local Channel Traffic.  }
   \label{fig:syn-3app-lc-traffic}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:samg-gc-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:syn-mg-gc-traffic}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:syn-cr-gc-traffic}
    \end{subfigure}%
   \caption{Each Application Router Global Channel Traffic.}
   \label{fig:syn-3app-gc-traffic}
\end{figure*}

\textbf{We can add another set of synthetic workload, in which three applications have similar data transfer amount, if necessary. }

@@

\section{Conclusion}
The conclusion goes here.

>>>>

\subsection{Simulation Tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability\cite{codes}.  CODES support dragonfly network with high fidelity packet-level simulation\cite{codes-dragonfly}. It has this network workload component that is capable of taking real MPI application traces generated by SST DUMPI\cite{sst} for trace-driven simulations.

@@

Figure \ref{fig:3app-data-amount} shows the data transfer amount of each MPI rank of three application. Based on this figure, we could say that CrystalRouter is the most communication-intensive and AMG is the lest communication-intensive.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{data_amount}
  \caption{Applications Data Transfer Amount. AMG has the least amount of data transfer, compared with CR and MG.}
  \label{fig:3app-data-amount}
\end{figure}

In our experiments, the workload running on the dragonfly network under 6 different job placement and routine scheme configurations as listed in Table \ref{tab: placement routing configs}. We focus on the analysis of network status while the workload is running, as well as the performance of each individual application. We collected the following result data from the experiments:
\begin{itemize}
    \item router terminal link traffic and busy time.
    \item router local channel traffic and busy time.
    \item router global channel traffic and busy time.
    \item The time application spend on communication. 
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{wkld/wkld-commtime}
  \caption{Time spent on communication by Workload I when 6 different placement and routing scheme configurations are used. Random is better than contiguous in terms of improving communication efficiency.Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
  \label{fig:wkld-commtime}
\end{figure}

@@

Random placement can randomly allocate the MPI ranks to the physical nodes all over the network, thus prevent the traffic from being congested certain part of the network. Shown as the solid lines in Figure \ref{fig:global-channel-traffic}, the traffic go through global channels are quite uniformly distributed across all the routers in the network, thus no particular global channel will suffer high traffic burden. When the random placement coupled with minimal routing, the packets can avoid unnecessary intermediate forwarding as introduced by adaptive and progressive routing, which results in generating less traffic. That's why the blue solid line is lower than the red and green solid line, shown in figure\ref{fig:global-channel-traffic}. Since the random placement can uniformly distribute the workload traffic, there are no sudden surge of busy time of global channel over the network, shown as solid lines in Fgiure \ref{fig:global-channel-stime}.

@@

However, AMG is quite an exception. As shown in Figure \ref{fig:amg-commtime}, the communication time of AMG is skyrocketing when it running under random placement coupled with (progressive) adaptive routing. The contiguous placement coupled with (progressive) adaptive routing can reach the best performance for AMG. The performance of AMG running under contiguous placement coupled with minimal routing is no better than random coupled with minimal routing, but still much better than random with (progressive) adaptive routing.

% Show the evidence from network level, explain why AMG is different. 
We look into network level to identify the major culprit behind AMG's abnormal behavior with random placement and adaptive routing. By identifying the terminal nodes and routers that each application's traffic go through, we can see that the traffic amount go through the routers that belongs to AMG is skyrocketing when it running with random placement and adaptive routing, shown in Figure \ref{fig:amg-lc-traffic} and Figure \ref{fig:amg-gc-traffic}. Adaptive routing will redirect the traffic of other applications to go through the routers belongs to AMG, and increase the traffic burden on those routers, thus slow down the communication of AMG.

No router will be shared between applications under contiguous placement. A set of successive groups will be allocated to each application under contiguous placement scheme. When contiguous placement coupled with minimal routing, application will be free of interference from its running neighbors. However, application communication traffic will be confined within their allocated routers, thus local congestion is very likely to happen depends on application communication intensity. As the lest communication intensive application, AMG can benefit from contiguous placement. Under contiguous placement, the traffic go through the local and global channels of the routers that belongs to AMG is significantly lower than random placement, shown in Figure \ref{fig:amg-lc-traffic}, \ref{fig:amg-gc-traffic}. MultiGrid and CrystalRouter are more communication intensive compared with AMG. They will suffer great amount local traffic in both local and global channels when contiguous placement is used, shown as the dash line in Figure \ref{fig:mg-lc-traffic}, \ref{fig:mg-gc-traffic} and Figure \ref{fig:cr-lc-traffic}, \ref{fig:cr-gc-traffic}. The random placement scheme can uniformly distribute their traffic all over the network by allocating randomly selected nodes to each MPI rank.

@@

The reason for AMG being the only victim from random placement and adaptive routing is due to  its relatively small amount of data transfer, that make it the least communication-intensive application in the workload.

From the detailed analysis of three applications performance above, we know that random placement and adaptive routing is quite efficient for orchestrate the traffic load in the network. Random placement and adaptive routing can uniformly distribute the workload traffic all over the network, make the load-balanced and avoid hot-spots. However, application like AMG will be impaired in this orchestration. The routers through which the less communication-intensive application traffic used to route, now have more traffic go through them. These extra traffic are from other communication-intensive applications. Beneath the facade of harmony, less communication-intensive application are ``bullied" by their communication-intensive peers in the workload.

We have tried three different congestion sensing schemes that used in adaptive routing\cite{won-prog-adaptive}, although there are some variations of the experiments, none of the congestion sensing scheme can prevent the ``bully" from happening.

@@

In order to prove that AMG being bullied only because it is less communication-intensive than its running peers, we generate a synthetic application, sAMG, which are same the with AMG's original traces in every aspect expect that the data transfer amount 100x greater than AMG. We run the new workload that consists of sAMG, MultiGrid and CrystalRouter on the same dragonfly network with 6 different placement and routing configurations. As shown in Figure \ref{fig:syn-3app-data-amount}, sAMG become the most communication-intensive application in the new workload.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{syn-wkld/data_amount}
  \caption{Applications Data Transfer Amount. sAMG has the most amount of data transfer, compared with CR and MG.}
  \label{fig:syn-3app-data-amount}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{syn-wkld/wkld-commtime}
  \caption{ Time spent on communication by Workload II when 6 different placement and routing scheme configurations are used. Random is better than contiguous in terms of improving communication efficiency.Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
  \label{fig:syn-wkld-commtime}
\end{figure}

\TODO{Explain} Figure \ref{fig:syn-wkld-commtime}.

The communication time of each application shown in Figure \ref{fig:syn-apps-commtime}. The ``bully" become the ``bullyee". The less communication-intensive applications in the workload this time are  MultiGrid and CrystalRouter.  When random placement coupled with (progressive) adaptive routing is used, MultiGrid and CrystalRouter suffer prolonged communication time, shown in Figure \ref{fig:syn-mg-commtime}, \ref{fig:syn-cr-commtime}. sAMG on the other hand benefit from random placement, shown in Figure \ref{fig:syn-samg-commtime }.

Contiguous placement coupled with minimal routing prevent applications from sharing network resource, and confine application communication traffic to their allocated routers. This can guarantee the performance of less communication intensive applications like MultiGrid and CrystalRouter, and make the communication intensive application like sAMG severely deteriorate.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/commtime}
        \caption{sAMG Communication Time}
        \label{fig:syn-samg-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/commtime}
        \caption{MG Communication Time}
        \label{fig:syn-mg-commtime}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/commtime}
        \caption{CR Communication Time}
        \label{fig:syn-cr-commtime}
    \end{subfigure}%
   \caption{Application communication time. Three applications running concurrently on dragonfly network with different placement and routing configurations. Random placement and adaptive routing can improves AMG's communication time. MG and CR communication time are prolonged with random placement and adaptive routing.}
   \label{fig:syn-apps-commtime}
\end{figure*}

% explain the network status of sAMG, MG and CR
We look into the network level to scrutinize the traffic went through the routers belongs to each application. 
Figure \ref{fig:syn-samg-lc-traffic}, \ref{fig:syn-samg-gc-traffic} shows the traffic go through the local and global channels of the routers that belongs to sAMG when different placement and routing configuration is used. The traffic is skyrocketing in both local and global channels when contiguous placement coupled with minimal routing. The contiguous placement make all the MPI ranks of sAMG reside on a set of nodes in the groups that adjacent to each other and minimal routing confines the traffic in those groups, results in severe local congestion. When adaptive routing is used, the local congestion will be alleviated in both local and global channels, as some traffic will be redirected to other part of the network. Random placement scheme can further reduce the traffic burden on local and global channels by assigning each MPI rank of sAMG with randomly select nodes, thus uniformly distribute sAMG traffic over the network.

MultiGrid become a victim of sAMG's bully when running under random placement. The traffic go through the local and global channels of the routers that belongs to MultiGrid are shown in Figure \ref{fig:syn-mg-lc-traffic}, \ref{fig:syn-mg-gc-traffic}. Although contiguous placement allocates all the MPI ranks in set of nodes in adjacent groups, causing possible local congestion, it will save MultiGrid's routers being shared by the traffic from its communication intensive peer, sAMG. As random placement is used, the traffic go through both local and global channels are greatly increased.

CrystalRouter has more traffic than MultiGrid. The random placement won't redirect much traffic from other applications to the routers belongs to CrystalRouter, compared with contiguous placement, as shown in Figure \ref{fig:syn-cr-lc-traffic}, \ref{fig:syn-cr-gc-traffic}.

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/lc-traffic}
        \caption{sAMG Local Channel Traffic}
        \label{fig:syn-samg-lc-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:syn-mg-lc-traffic}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:syn-cr-lc-traffic}
    \end{subfigure}%
   \caption{Each Application Router Local Channel Traffic.  }
   \label{fig:syn-3app-lc-traffic}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/gc-traffic}
        \caption{sAMG Global Channel Traffic}
        \label{fig:syn-samg-gc-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:syn-mg-gc-traffic}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:syn-cr-gc-traffic}
    \end{subfigure}%
   \caption{Each Application Router Global Channel Traffic.}
   \label{fig:syn-3app-gc-traffic}
\end{figure*}

@@

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have conducted extensive study about application's behavior when it running on dragonfly network with different job placement and routing configurations. We resorted to CODES, a HPC network simulation tool with high fidelity, and used three parallel applications traces collected from production system to analysis the applications' behavior on network level. We found that when applications are running concurrently, although random placement can uniformly distribute the workload traffic over the network, alleviate local congestion, the performance of certain application would be impaired. We identify this phenomenon as ``bully" between concurrently running jobs, and the victim always being the less communication-intensive application. On the other hand, the contiguous placement tend to cause local congestion, however when coupled with adaptive routing, the configuration can guarantee the performance of ``bullee" application by limiting the network sharing between concurrently running application.

Due the presence of this ``bully", we suggest that when the batch scheduler make job placement for workload on system with dragonfly network, random placement shouldn't be the rule of thumb. Instead, the placement should made with consideration of jobs communication intensity. The less intensive application should get contiguous allocation coupled with adaptive routing, the intensive ones should get random allocation coupled with minimal routing.

We believe the findings presented in this paper can illuminate a path to more efficient workload manager for future systems with dragonfly network.
